---
title: "Untitled"
output:
  word_document: default
  html_document: default
date: "2023-12-08"
---

```{r }
df_glass =read.table('glass_data.txt', header = FALSE, sep = ',', na.strings = '.')
library(ggplot2)
```

```{r }
colnames(df_glass) = c('Id number',  'RI', 'Na','Mg', 'Al', 'Si', 'K_Pota' , 'Ca', 'Ba','Fe',  'glass_type')
```

```{r }
summary(df_glass)
```
```{r}
X <- df_glass[,1:9]
par(mfrow = c(3, 3))
for (i in 1:ncol(X)) {
  hist(X[ ,i], xlab = names(X[i]), main = paste(names(X[i]), "Histogram"), col="steelblue")  
}
```
```{r }
pairs(X, main="Scatterplot Matrix", col='steelblue') 

```
```{r }
X <- df_glass[,1:9]
par(mfrow = c(3, 3))
for (i in 1:ncol(X)) {
  boxplot(X[ ,i], ylab = names(X[i]), horizontal=T,
          main = paste(names(X[i]), "Boxplot"), col="steelblue")
}

```
```{r }
par(mfrow = c(3, 3))
for (i in 1:ncol(X)) {
  d <- density(X[,i], na.rm = TRUE)
  plot(d, main = paste(names(X[i]), "Density"))
  polygon(d, col="#009999")
}

``` 
```{r }
X <- df_glass[, 1:9]

# Setting up the layout for multiple plots
par(mfrow = c(3, 3))

for (i in 1:ncol(X)) {
  if (is.numeric(X[, i])) {
    # Compute average values by 'glass_type'
    avg_values <- aggregate(X[, i], by = list(glass_type = df_glass$glass_type), FUN = mean)
    
    # Bar plot for each numeric variable against 'glass_type'
    barplot(avg_values$x, names.arg = avg_values$glass_type, 
            xlab = "Glass Type", ylab = paste("Average", names(X)[i]), 
            col = "#009999", main = paste("Average", names(X)[i], "by Glass Type"))
  }
}

```

```{r}
#df_glass <- df_glass[, -1] # Remove the first column by index
df_glass <- df_glass[, -which(names(df_glass) == "Id number")] # Remove by column name
```

```{r fig.width=14, warning=FALSE}
library(GGally)
pairs_plot <- GGally::ggpairs(df_glass, 
  aes(color = ""), 
  lower = list(continuous = "points"),
  upper = list(continuous = "cor"),
  diag = list(continuous = "barDiag")
)
pairs_plot <- pairs_plot + scale_color_manual(values = "darkgreen") 
print(pairs_plot)

```
for Caption : Ca (Calcium) and RI (refractive index) are rather highly correlated (positively). From the scatter plots here below it seems that the relationship between is linear, RI increases as Ca increases.


```{r }
correlation_matrix <- cor(df_glass)
# Plotting the heatmap
#ggcorrplot(correlation_matrix, hc.order = TRUE, type = "lower", lab = TRUE)
```
```{r }
model_type <- lm(glass_type ~ Na + Mg + Al + Si + K_Pota + Ca + Ba + Fe, data = df_glass)
summary(model_type )
```   
```{r }
par(mfrow=c(2,2))
plot (model_type)
```

```{r}
model_ri <- lm(RI ~ Na + Mg + Al + Si + K_Pota + Ca + Ba + Fe, data = df_glass)
summary(model_ri)
```

```{r }
model_ri <- lm(glass_type ~ Na + Mg + Al + Si + K_Pota + Ca + Ba + Fe, data = df_glass)
summary(model_ri)

```

```{r}

ggplot(df_glass, aes(x = factor(glass_type), y = Al, fill = factor(glass_type))) +
  geom_boxplot() +
  labs(
    x = "Glass Type",
    y = "Al",
    fill = "Glass Type"
  )
ggplot(df_glass, aes(x = factor(glass_type), y = Na, fill = factor(glass_type))) +
  geom_boxplot() +
  labs(
    x = "Glass Type",
    y = "Na",
    fill = "Glass Type"
  )

```

**Plot  for signifcabt variable for RI**
```{r }
# Assuming df_glass is your dataframe

par(mfrow = c(2, 2))  # Setting up the layout for multiple plots

# Scatterplot and trend line for 'Na' vs 'RI'
plot(df_glass$Na, df_glass$RI, 
     xlab = "Na", ylab = "RI", 
     main = "RI vs Na", col = "#009999")
abline(lm(RI ~ Na, data = df_glass), col = "red")

# Scatterplot and trend line for 'Mg' vs 'RI'
plot(df_glass$Mg, df_glass$RI, 
     xlab = "Mg", ylab = "RI", 
     main = "RI vs Mg", col = "#009999")
abline(lm(RI ~ Mg, data = df_glass), col = "red")

# Scatterplot and trend line for 'K_Pota' vs 'RI'
plot(df_glass$K_Pota, df_glass$RI, 
     xlab = "K_Pota", ylab = "RI", 
     main = "RI vs K_Pota", col = "#009999")
abline(lm(RI ~ K_Pota, data = df_glass), col = "red")

# Scatterplot and trend line for 'Ca' vs 'RI'
plot(df_glass$Ca, df_glass$RI, 
     xlab = "Ca", ylab = "RI", 
     main = "RI vs Ca", col = "#009999")
abline(lm(RI ~ Ca, data = df_glass), col = "red")

```
```{r fig.width=10}
X <- df_glass[, 2:(ncol(df_glass) - 1)]  # Excluding the 'RI' and 'glass_type' variables

# Setting up the layout for multiple plots
par(mfrow = c(3, 3))

for (i in 1:ncol(X)) {
  if (is.numeric(X[, i])) {
    # Density plot for each variable against 'RI'
    density_values <- density(X[, i])
    
    plot(density_values, type = 'l', 
         xlab = "RI", ylab = "Density",
         col = "#009999", main = paste(names(X)[i], "Density vs RI"))
    polygon(density_values, col = adjustcolor("#009999", alpha.f = 0.3), border = NA)
  }
}


``` 
```{r }
X <- df_glass[, 2:(ncol(df_glass) - 1)]  # Excluding the 'RI' and 'glass_type' variables

# Calculate correlation matrix
correlation_matrix <- cor(cbind(X, RI = df_glass$RI))

# Create a heatmap of correlations
library(ggplot2)
library(reshape2)

correlation_df <- melt(correlation_matrix)
ggplot(correlation_df, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = '#009999') +
  theme_minimal() +
  labs(title = "Correlation Heatmap of RI with Other Variables")


```
```{r fig.width=10}
X <- df_glass[, 2:(ncol(df_glass) - 1)]  # Excluding 'RI' and 'glass_type'

# Setting up the layout for multiple plots
par(mfrow = c(3, 3))

for (i in 1:ncol(X)) {
  plot(df_glass$RI, X[, i], 
       xlab = "RI", ylab = names(X)[i],
       col = "#009999", main = paste("RI vs", names(X)[i]))
}



```

```{r }
# Subset the data excluding 'RI' and 'glass_type'
X <- df_glass[, 2:(ncol(df_glass) - 1)]

# Calculate correlation matrix
correlation_matrix <- cor(cbind(X, RI = df_glass$RI))

# Plotting heatmap using the 'corrplot' package
library(corrplot)

# Displaying the correlation matrix as a heatmap
corrplot(correlation_matrix, method = "color", type = "upper",
         addCoef.col = "black", tl.col = "black",
         title = "Correlation Heatmap of RI with Other Variables")



```

```{r}
ggplot(df_glass, aes(x = Na, y = RI, color = RI)) +
  geom_point() +
  labs(
    x = "Na",
    y = "RI",
    color = "RI"
  ) +
  theme_minimal()

```

```{r }
library(ggplot2)

# Assuming your data frame is df_glass
ggplot(df_glass, aes(x = factor(glass_type), y = Al)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "Glass Type", y = "Al") +
  ggtitle("Boxplot of Al by Glass Type")
 

```
```{r }
# Assuming your data frame is df_glass
ggplot(df_glass, aes(x = factor(glass_type), y = Na)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "Glass Type", y = "Na") +
  ggtitle("Boxplot of Al by Glass Type")

```
```{r }
# Assuming your data frame is df_glass
ggplot(df_glass, aes(x = factor(glass_type), y = Na)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "Glass Type", y = "Al") +
  ggtitle("Boxplot of Al by Glass Type")


```

```{r }
# Assuming your data frame is df_glass
ggplot(df_glass, aes(x = factor(glass_type), y = Al, color = factor(glass_type))) +
  geom_point(position = position_jitter(width = 0.3), alpha = 0.6, size = 3) +
  labs(x = "Glass Type", y = "Al") +
  ggtitle("Scatter Plot of Al by Glass Type") +
  scale_color_discrete(name = "Glass Type")

```

```{r }
df_glass$glass_type=as.factor(df_glass$glass_type)
str(df_glass)
```
```{r}


```

```{r}
library(ggplot2)
plot <- ggplot(df_glass, aes(x = RI, y = Na, color = type)) +
  geom_point() +
  labs(
    x = "RI",
    y = "Na",
    color = "Type",
    title = "Scatter Plot of RI vs Na with Type",
    caption = 'Relationship of Sodium (NA) with Reflexive Index and Galss Type')
# Apply theme to modify caption appearance
plot + theme(
  plot.caption = element_text(
    color = "steelblue",                     # Change text color
    size = ,                          # Adjust text size
    hjust = 0.5,                        # Horizontal alignment
    margin = margin(t = 10, l = 20, unit = "pt"), # Add margin from the top
    #position= 'top'
  )
)
#Na is an important feature for glass type
#Ri distribution does not provide much insight


```
```{r }



plot <- ggplot(df_glass, aes(x = Al, y = RI, color = type)) +
  geom_point() +
  labs(
    x = "Al",
    y = "RI",
    color = "Type",
    title = "Scatter Plot of RI vs Na with Type",
    caption = 'Relationship of Sodium (NA) with Reflexive Index and Galss Type')
# Apply theme to modify caption appearance
plot + theme(
  plot.caption = element_text(
    color = "steelblue",                     # Change text color
    size = ,                          # Adjust text size
    hjust = 0.5,                        # Horizontal alignment
    margin = margin(t = 10, l = 20, unit = "pt"), # Add margin from the top
    #position= 'top'
  )
)
```

```{r}
plot <- ggplot(df_glass, aes(x = Mg, y = Al, color = type)) +
  geom_point() +
  labs(
    x = "Al",
    y = "Na",
    color = "Type",
    title = "Scatter Plot of NA vs Mg with Type",
    caption = ''
  )

# Apply theme to modify caption appearance
plot + theme(
  plot.caption = element_text(
    color = "black",                     # Change text color
    size = 10,                          # Adjust text size
    hjust = 0.5,                        # Horizontal alignment
    margin = margin(t = 15, unit = "pt") # Add margin from the top
  )
)

